Accurate evaluation of dislocation tractions

### 2023-04-20 src/wip/DanielCelisGarza/Accurate_evaluation_of_dislocation_tractions/paper/main.tex
Commiting before final upload to ora

In simulations, tractions are manifested as image stresses calculated by the FE solver at FE nodes. In order to validate and compare the practical differences between analytic and numeric methods, we compare the resulting image stresses from both methods to the analytic expressions for infinite dislocations in inhomogenous media for edge dislocations \cite{head1953edge}, as well as those for screw dislocations \cite[p.~59,~64]{hirth1983theory}. We keep the same nomenclature and coordinate system as both infinite-domain solutions. Where the traction surface is the line $x=0$, the dislocation line direction is the positive $z$-direction (pointing out of the page), the dislocation coordinates are represented by $(a,~c)$, and points in the $xy$-plane described by their $(x,\,y)$ coordinates.

@@

Note the first terms of \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew} all correspond to the stress field generated by the dislocation itself. The following terms are the corrective terms required to make the boundary conditions on the surface equal to zero. Therefore, we can split these equations and only look at the real or corrective terms independently, which we do in order to only visualise the effect of the different traction calculations. Furthermore, \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew} are all singular at the dislocation coordinates. Our simulation code uses the non singular expressions found by \citet{Cai2006}, which smooth out drastic increases in stresses and avoid numerical blow up as we near the dislocation core.

@@

\citet{Queyreau} found the analytic solution is approximately 10 times more computationally expensive than its numerical counterpart for 1 quadrature point, our findings agree with this result but the implications for a whole simulation are favourable (see \cref{s:results}).

@@

Furthermore, the chirality and self-consistency of the FE nodes must be accounted for such that they are in the proper order regardless of the element face they belong to. Here we use 8-node linear hexahedral (brick) elements. The node ordering for the various surfaces is that for which the calculated normals point out from the the domain, this is dependant on the specific FE mesh implementation.

@@

\Cref{a:tot_force} illustrates how the total force on each node is obtained. However, our implementation does not strictly follow it because we memoise a generalised version of $\vec{L}$ upon simulation initialisation instead of finding one at every iteration, reducing computational time but requiring us to account for nodes without traction boundary conditions. Our indexing also starts at 1, but zero indexing makes the algorithm easier to follow.
\begin{algorithm}
    \caption{Assuming $ \vec{\tilde{F}} $ is arranged the same way as $ \vec{\gamma} $ and indexing starts at 0.}
    \begin{algorithmic}[1]
        \label{a:tot_force}
        \State\Comment{Loop through the array containing the node labels of the relevant surface nodes.}
        \For{$ i = 0;\, i < \rvar{length}(\vec{\gamma});\, i++$}					\State\Comment{Save the global node label for the current iteration.}
        \State $ n \gets \vec{\gamma}[i] $
        \State\Comment{Use the node label to find a vector, $\vec{L}$, with the linearised indices in $\mtx{N_{L}}$ where node $n$ appears as part of a surface element whose tractions we are calculating.}
        \State $ \vec{L} \gets \rvar{find}(\mtx{N_{L}} == n) $
        \State\Comment{Loop over coordinates.}
        \For{$ k = 0;\, k < 3;\, k++ $}
        \State\Comment{Use global node label vector to index the force array from the analytical force calculation. Multiplied by 3 because there are three coordinates per node. We sum the forces from the analytical calculation because the same global node can be part of multiple surface elements. We add $ k $ because the $ x,~y,~z $ coordinates are consecutively stored in $ \mtx{F_{e}} $.}
        \State $ \vec{\tilde{F}}[3n + k] \gets \vec{\tilde{F}}[3n + k] + \sum\mtx{F_{e}}[3\vec{L} + k]  $
        \EndFor
        \EndFor
    \end{algorithmic}
\end{algorithm}

@@

The limitations become even more evident when the dislocation line segment is parallel to a surface element. In \cref{f:rel_err_par_edge}, we observe the relative errors are quite large when close to the surface. At distances larger than $10^4\, $, loss of significance causes the relative errors to converge at $\sim 1$ which is expected as two finite precision floating point numbers get closer to zero. Of particular note is how large the relative errors are even at $100$ lattice units away from the surface, even for large numbers of quadrature points. This figure backs up the earlier point regarding Gauss nodes close to maximal values of rational functions. It can be shocking to see that as $Q = 2$ performs significantly better than $Q = 10, 11, 100, 100$ at distances from the surface as having nodes near the maximal values is a large source of error. Consequently, different configurations have different optimal numbers of points. The numerical instability of this method, which when coupled to the chaotic nature of dislocation dynamics and stiffness of the equations of motion, can lead to large deviations between simulations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{par_e_z.pdf}
    \caption{Log-log plot of the relative error as a function of distance from the surface element for a parallel edge dislocation (\cref{f:gauss_quad_test}b). Burgers vector, $\vec{b} = [0\, 1\, 0]$, line direction, $\vec{l} = [1\, 0\, 0]$, dislocation core and surface element parameters are the same as \cref{f:rel_err_perp_edge}. The dislocation length is fixed to $10^{6},\, x\in\left[-0.5\times10^{6},\, 0.5\times10^{6} \right]$ and bisects the surface element along the $[1\, 0\, 0]$ direction. The whole dislocation was segmented into $10^4$ pieces of length $100$ to prevent the dislocation from intersecting the surface element when they were rotated to avoid the singularity. The relative error at large distances ($>10^6\, $) converges to one due to loss of significance.}
    \label{f:rel_err_par_edge}
\end{figure}

From \cref{f:rel_err_perp_edge,f:rel_err_par_edge} one might be tempted to say that for a segment parallel to a surface, Gauss quadrature performs far worse than for a perpendicular one. However there is a further wrinkle in this problem: symmetry. To exemplify this we plot the relevant components of the stress tensor for the arrangement found in \cref{f:gauss_quad_test}(a) in \cref{sf:xzeperp,sf:yzeperp,sf:zzeperp}. The $\sigma_{xz}$ and $\sigma_{zz}$ components are antisymmetric about the centre of the element. If we use Gauss quadrature on them, we sample equivalent but oppositely valued points that are equally weighted, thus the sum vanishes and therefore do not contribute to \cref{f:rel_err_perp_edge}. However, $\sigma_{yz}$ does not vanish, but can be accurately integrated with sufficiently large $Q$. If we were to move the dislocation off-centre such that these symmetries are broken, the errors would increase.

We also plot the relevant stresses for the arrangement described by \cref{f:gauss_quad_test}(b) in \cref{sf:xzepar,sf:yzepar,sf:zzepar}. From \cref{sf:xzepar} and \cref{sf:yzepar}, it is immediately apparent why gauss quadrature fails so spectacularly in \cref{f:rel_err_par_edge}. At low numbers of quadrature points, it fails to appropriately sample the rapidly changing value of $\tilde{\sigma}_{xz}$ at both ends of the dislocation, as well those in the neighbourhood of the dislocation in $\tilde{\sigma}_{yz}$. Moreover, the further away the quadrature points are from the midpoint of the domain, the lower their relative weighting. So even with a relatively large number of them, there can still be large errors. Which is a particularly egregious problem in \cref{sf:yzepar}, and explains why such a large number of quadrature points is required to accurately compute the integral.
\begin{figure}
    \centering
    \subfloat[\label{sf:xzeperp}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{sxzperpEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:yzeperp}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{syzperpEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:zzeperp}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{szzperpEdgeFiveb.pdf}}
    }

@@

Despite these being artificially idealised examples that illustrate the failings of Gauss quadrature, other problematic scenarios commonly show up in simulations. These tend to worsen with smaller core radii $a$, fewer Gauss nodes, higher dislocation densities near surfaces, more permissive mobility functions, and coarser FE meshes. The $\mathcal{O}(1/R)$ decay rate of $\tns{\tilde{\sigma}}$ and chaotic nature of dislocation dynamics, means these errors may result in unwarranted topological changes that cascade as the simulation advances. This is particularly deleterious when doing simulations with higher dislocation densities and/or where a large number of dislocations are close to the surface, such as nanoindentation simulations.

@@

    \caption{Image stresses for an edge dislocation with $\vec{l} = [0\, 0\, 1]$, $\vec{b} = [0\, 1\, 0]$, where $a = 5$, with coordinates $(26,\, 1000)\, $, i.e. the centre of the first FE from the surface at $x=0$, and in the centre of the simulation box from top to bottom. (a), (e) and (i) are the stress fields for the infinite-domain solution, $\tns{\hat{\sigma}}$; (b), (f) and (j) are those obtaine from analytic tractions + FEM, $\tns{\hat{\sigma}}^{\textrm{A}}$; (c), (g) and (k) are those obtained using numeric tractions ($Q = 1$) + FEM, $\tns{\hat{\sigma}}^{\textrm{N}}$. (a) to (c) represent the $xx$; (e) to (g) the $yy$; and (i) to (k) the $xy$ components of the stress tensor.}
    \label{f:head_vs_ana_vs_num_epar}
\end{figure}
\Cref{f:head_vs_ana_vs_num_eperp} shows the stress fields corresponding to analytic expressions for image stress components, $\hat{\sigma}_{ij}$ in \cref{eq:imageStressAnalyticEdge1} where no superscript denotes the infinite-domain singular expressions, the $\textrm{A}$ superscript are the stresses calculated from analytic tractions and the $\textrm{N}$ superscript are those coming from numeric tractions where $Q = 1$ (same nomenclature in \cref{f:head_vs_ana_vs_num_epar,f:head_vs_ana_vs_num_screw}). The setup corresponds to the one described in \cref{f:headvstractionfem} where $\vec{b} = \vec{b}_{\textrm{e1}} = [1\, 0\, 0]$ and the dislocation is found at $(26, 1000)$.

@@

    \subfloat[%$\hat{\sigma}_{yz}$.
        \hspace{-0.8cm}\label{sf:szyscrew}]
    {
        {\includegraphics[width=0.25\linewidth]{syzscrew_leg.pdf}}
    }~
    \subfloat[%$\hat{\sigma}^\textrm{A}_{yz}$.
        \hspace{-0.8cm}\label{sf:syzAscrew}]
    {
        {\includegraphics[width=0.25\linewidth]{syzAscrew_leg.pdf}}
    }~
    \subfloat[%$\hat{\sigma}^\textrm{N}_{yz}$.
        \hspace{-0.8cm}\label{sf:syzNscrew}]
    {
        {\includegraphics[width=0.25\linewidth]{syzNscrew_leg}}
    }~
    \subfloat{
        \includegraphics[height = 0.227\linewidth]{syzNAscrew_leg.pdf}
    }
    \caption{Image stresses for a screw dislocation with $\vec{l} = [0\, 0\, 1]$, $\vec{b} = [0\, 0\, 1]$, where $a = 5$, with coordinates $(26,\, 1000)\, $, i.e. the centre of the first FE from the surface at $x=0$, and in the centre of the simulation box from top to bottom.  (a) and (e) are the stress fields for the infinite-domain solution, $\tns{\hat{\sigma}}$; (b) and (f) are those obtaine from analytic tractions + FEM, $\tns{\hat{\sigma}}^{\textrm{A}}$; (c) and (g) are those obtained using numeric tractions ($Q = 1$) + FEM, $\tns{\hat{\sigma}}^{\textrm{N}}$. (a) to (c) represent the $xz$; (e) to (g) the $yz$.}
    \label{f:head_vs_ana_vs_num_screw}
\end{figure}
Things get markedly more interesting when looking at $\vec{b} = \vec{b}_{\textrm{e2}} = [0\, 1\, 0]$ in \cref{f:head_vs_ana_vs_num_epar} for a dislocation in the same place, $(26, 1000)$. Of particular note is $\hat{\sigma}_{xx}$, where a comparison between \cref{sf:sxxepar,sf:sxxAepar} and \cref{sf:sxxNepar} reveals one of the major issues with numeric tractions. If we look at the neighbourhood of the dislocation (just to the right), we will find a sign inversion i.e. yellow and green contours as opposed to purple and blue. As well as a drastically different isosurface shape. Image stresses like those can lead dislocations to behave quite differently than they should, particularly if the sign inversion also has a significantly different magnitude than the correct solution. Specifically, there is a region in the positive $x$-direction away from the dislocation where $\hat{\sigma}_{xx}$ is tensile rather than compressive. It is as absurd as a ship that floats by lowering the water level.

@@

    \subfloat[$\hat{\sigma}_{xz}$.\label{sf:line_sxzscrew}]
    {
        {\includegraphics[width=0.3\linewidth]{line_sxzscrew_leg.pdf}}
    }~
    \subfloat[$\hat{\sigma}_{yz}$.\label{sf:line_syzscrew}]
    {
        {\includegraphics[width=0.3\linewidth]{line_syzscrew_leg.pdf}}
    }
    \caption{Line plots of the image stresses of a dislocation at $(26, 1000)$ for a line going through $x = 103$ (start of the third element) for the analytic image stresses, as well as those calculated with numeric and analytic tractions. (a) to (c) correspond to $\vec{b} = \vec{b}_{\textrm{e1}}$, (d) to (f) $\vec{b} = \vec{b}_{\textrm{e2}}$ and (g) to (h) to $\vec{b} = \vec{b}_{\textrm{s}}$.}
    \label{f:line_head_vs_ana_vs_num}
\end{figure}
From \cref{f:head_vs_ana_vs_num_eperp,f:head_vs_ana_vs_num_epar,f:head_vs_ana_vs_num_screw} it seems like both analytic and numeric tractions are appropriate in most cases. At least at these scales, there is only one instance where numeric tractions yield very incorrect results. That said, these can have a significant impact on simulations, particularly those where multiple dislocations interact with surfaces in close proximity with one another.

@@

From \cref{f:head_vs_ana_vs_num_epar2to11,f:line_head_vs_ana_vs_num_epar2to11} it can be observed that both analytic and numeric tractions converge to similar shapes to each other as expected. However it also becomes clear that the infinite-domain solution is not totally accurate for image stresses in finite domains due to edge effects.

Lastly, to show how much numeric tractions can affect a result, \cref{f:simulation} shows a few snapshots of the simulation described in \cref{s:method}. The figures shown are not at equivalent times because the simulations using numeric tractions ran indefenitely. The line tension equilibrated with the image forces on the traction-free surface, preventing the dislocation from exiting and keeping it oscillating in a local energy minimum. The simulation using analytic tractions ended when the dislocation completely exited the surface as expected. It is feasable that at some point, the simulation using numeric tractions could jump out of the local minimum due to some spike in the tractions, but the last image was taken for a simulation time approximately 20 times greater than it took the one using analytic tractions to end when the dislocation fully exited the box.
\begin{figure}
    \subfloat[Analytic tractions $t = 0$.]
    {
        {\includegraphics[width=0.3\linewidth]{analytic_0.pdf}}
    }~
    \subfloat[Analytic tractions mid simulation.]
    {
        {\includegraphics[width=0.3\linewidth]{analytic_50.pdf}}
    }~
    \subfloat[Analytic tractions end.]
    {
        {\includegraphics[width=0.3\linewidth]{analytic_66.pdf}}
    }

    \subfloat[Numeric tractions $t = 0$.]
    {
        {\includegraphics[width=0.3\linewidth]{numeric_0.pdf}}
    }~
    \subfloat[Numeric tractions mid simulation.]
    {
        {\includegraphics[width=0.3\linewidth]{numeric_78.pdf}}
    }~
    \subfloat[Numeric tractions $t \to \infty$.]
    {
        {\includegraphics[width=0.3\linewidth]{numeric_156.pdf}}
    }
    \caption{Progression of simulations using both traction calculation methods.}
    \label{f:simulation}
\end{figure}

@@

Often the effects of tractions on a simulation can be quite subtle if things do not go catastrophically wrong, but a quick and easy way of seeing how the numeric tractions have a tendancy to spike and invert sign (as in \cref{sf:line_sxxepar}) is to pause a simulation while its running and scatter plot the numeric tractions vs analytic ones. The plot will show a positive correlation---as the sign inversions in numeric tractions are relatively rare events---but whichever axis corresponds to the numeric tractions will have the largest range. Using proportional axes makes the differences quite evident even at a quick glance as the aspect ratio will be far from 1:1.

In other parts of our work, we have noted that despite the analytic tractions taking approximately 10 times longer to compute than numerically calculated tractions with $Q = 1$ (which agrees with the findings in \cite{Queyreau}), yield more stable and ultimately faster simulations. Even simple simulations are typically faster when using analytic tractions over numeric ones. This was not the case with the simulation we show here, but the numeric tractions led to a hung simulation, which we have found is quite common. The margin by which using analytic tractions leads to faster overall simulations grows as simulation complexity increases. Only the simplest simulations initially benefit from numeric tractions but this often decreases and reverses as the simulation advances. Another fortunate side effect of more accurate tractions is the fact that fewer dislocation segments are generated during simulations, which \citet{bromage2018calculating} found when correctly accounting for the displacements generated by dislocations.

Given our findings, we cannot recommend using numeric tractions when analytic ones are available. The losses in computational speed that result from moving from one to the other are more than made up for by the fewer topological operations, fewer generated segments, and more accurate velocities. All of which result in fewer calculations of segment-segment interactions, fewer collisions, larger timesteps and ultimately cleaner simulations that run into fewer snags along the way.

>>>>

In simulations, tractions are manifested as image stresses calculated by the FE solver at FE nodes. In order to validate and compare the practical differences between analytic and numeric methods, we compare the resulting image stresses from both methods, to the analytic expressions for infinite dislocations in inhomogenous media for edge \cite{head1953edge}, and screw dislocations \cite[p.~59,~64]{hirth1983theory}. We keep the same nomenclature and coordinate system as both infinite-domain solutions. Where the traction surface is the line $x=0$, the dislocation line direction is the positive $z$-direction (pointing out of the page), the dislocation coordinates are represented by $(a,~c)$, and points in the $xy$-plane described by their $(x,\,y)$ coordinates.

@@

Note the first terms of \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew} all correspond to the stress field generated by the dislocation itself. The following terms are the corrective terms required to make the boundary conditions on the surface equal to zero. Therefore, we can split these equations and only look at the real or corrective terms independently, which we do in order to only visualise the effect of the different traction calculations. Furthermore, \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew} are all singular at the dislocation coordinates. Our simulation code uses the non-singular expressions found by \citet{Cai2006}, which smooth out drastic increases in stresses and avoid numerical blow up as we near the dislocation core.

@@

\citet{Queyreau} found the analytic solution is approximately 10 times more computationally expensive than its numerical counterpart for 1 quadrature point, our findings agree with this result, but the implications for a whole simulation are favourable (see \cref{s:results}).

@@

Furthermore, the chirality and self-consistency of the FE nodes must be accounted for such that they are in the proper order regardless of the element face they belong to. Here we use 8-node linear hexahedral (brick) elements. The node ordering for the various surfaces is that for which the calculated normals point out from the domain, this is dependent on the specific FE mesh implementation.

@@

\Cref{alg:tot_force} illustrates how the total force on each node is obtained. However, our implementation does not strictly follow it because we memoise a generalised version of $\vec{L}$ upon simulation initialisation instead of finding one at every iteration, reducing computational time but requiring us to account for nodes without traction boundary conditions. Our indexing also starts at 1, but zero indexing makes the algorithm easier to follow.
\begin{algorithm}
    \caption{Assuming $ \vec{\tilde{F}} $ is arranged the same way as $ \vec{\gamma} $ and indexing starts at 0.}
    \begin{algorithmic}[1]
        \label{alg:tot_force}
        \State\Comment{Loop through the array containing the node labels of the relevant surface nodes.}
        \For{$ i = 0;\, i < \rvar{length}(\vec{\gamma});\, i++$}					\State\Comment{Save the global node label for the current iteration.}
        \State $ n \gets \vec{\gamma}[i] $
        \State\Comment{Use the node label to find a vector, $\vec{L}$, with the linearised indices in $\mtx{N_{L}}$ where node $n$ appears as part of a surface element whose tractions we are calculating.}
        \State $ \vec{L} \gets \rvar{find}(\mtx{N_{L}} == n) $
        \State\Comment{Loop over coordinates.}
        \For{$ k = 0;\, k < 3;\, k++ $}
        \State\Comment{Use global node label vector to index the force array from the analytical force calculation. Multiplied by 3 because there are three coordinates per node. We sum the forces from the analytical calculation because the same global node can be part of multiple surface elements. We add $ k $ because the $ x,~y,~z $ coordinates are consecutively stored in $ \mtx{F_{e}} $.}
        \State $ \vec{\tilde{F}}[3n + k] \gets \vec{\tilde{F}}[3n + k] + \sum\mtx{F_{e}}[3\vec{L} + k]  $
        \EndFor
        \EndFor
    \end{algorithmic}
\end{algorithm}

@@

The limitations become even more evident when the dislocation line segment is parallel to a surface element. In \cref{f:rel_err_par_edge}, we observe the relative errors are quite large when close to the surface. At distances larger than $10^4$, loss of significance causes the relative errors to converge at $\sim 1$ which is expected as two finite precision floating point numbers get closer to zero. Of particular note is how large the relative errors are even at $100$ lattice units away from the surface, even for large numbers of quadrature points. This figure backs up the earlier point regarding Gauss nodes close to maximal values of rational functions. It can be shocking to see that as $Q = 2$ performs significantly better than $Q = 10, 11, 100, 100$ at distances from the surface as having nodes near the maximal values is a large source of error. Consequently, different configurations have different optimal numbers of points. The numerical instability of this method, which when coupled to the chaotic nature of dislocation dynamics and stiffness of the equations of motion, can lead to large deviations between simulations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{par_e_z.pdf}
    \caption{Log-log plot of the relative error as a function of distance from the surface element for a parallel edge dislocation (\cref{f:gauss_quad_test}b). Burgers vector, $\vec{b} = [0\, 1\, 0]$, line direction, $\vec{l} = [1\, 0\, 0]$, dislocation core and surface element parameters are the same as \cref{f:rel_err_perp_edge}. The dislocation length is fixed to $10^{6},\, x\in\left[-0.5\times10^{6},\, 0.5\times10^{6} \right]$ and bisects the surface element along the $[1\, 0\, 0]$ direction. The whole dislocation was segmented into $10^4$ pieces of length $100$ to prevent the dislocation from intersecting the surface element when they were rotated to avoid the singularity. The relative error at large distances ($>10^6$) converges to one due to loss of significance.}
    \label{f:rel_err_par_edge}
\end{figure}

From \cref{f:rel_err_perp_edge,f:rel_err_par_edge} one might be tempted to say that for a segment parallel to a surface, Gauss quadrature performs far worse than for a perpendicular one. However, there is a further wrinkle in this problem: symmetry. To exemplify this we plot the relevant components of the stress tensor for the arrangement found in \cref{f:gauss_quad_test}(a) in \cref{sf:xzeperp,sf:yzeperp,sf:zzeperp}. The $\sigma_{xz}$ and $\sigma_{zz}$ components are antisymmetric about the centre of the element. If we use Gauss quadrature on them, we sample equivalent but oppositely valued points that are equally weighted, thus the sum vanishes and therefore do not contribute to \cref{f:rel_err_perp_edge}. However, $\sigma_{yz}$ does not vanish, but can be accurately integrated with sufficiently large $Q$. If we were to move the dislocation off-centre such that these symmetries are broken, the errors would increase.

We also plot the relevant stresses for the arrangement described by \cref{f:gauss_quad_test}(b) in \cref{sf:xzepar,sf:yzepar,sf:zzepar}. From \cref{sf:xzepar} and \cref{sf:yzepar}, it is immediately apparent why gauss quadrature fails so spectacularly in \cref{f:rel_err_par_edge}. At low numbers of quadrature points, it fails to appropriately sample the rapidly changing value of $\tilde{\sigma}_{xz}$ at both ends of the dislocation, as well those in the neighbourhood of the dislocation in $\tilde{\sigma}_{yz}$. Moreover, the further away the quadrature points are from the midpoint of the domain, the lower their relative weighting. So even with a relatively large number of them, there can still be large errors. Which is a particularly egregious problem in \cref{sf:yzepar}, and explains why so many quadrature points are required to accurately compute the integral.
\begin{figure}
    \centering
    \subfloat[\label{sf:xzeperp}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{sxzperpEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:yzeperp}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{syzperpEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:zzeperp}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{szzperpEdgeFiveb.pdf}}
    }

@@

Despite these being artificially idealised examples that illustrate the failings of Gauss quadrature, other problematic scenarios commonly show up in simulations. These tend to worsen with smaller core radii $a$, fewer Gauss nodes, higher dislocation densities near surfaces, more permissive mobility functions, and coarser FE meshes. The $\mathcal{O}(1/R)$ decay rate of $\tns{\tilde{\sigma}}$ and chaotic nature of dislocation dynamics, means these errors may result in unwarranted topological changes that cascade as the simulation advances. This is particularly deleterious when doing simulations with higher dislocation densities and/or where numerous dislocations are close to the surface, such as nanoindentation simulations.

@@

    \caption{Image stresses for an edge dislocation with $\vec{l} = [0\, 0\, 1]$, $\vec{b} = [0\, 1\, 0]$, where $a = 5$, with coordinates $(26,\, 1000)$, i.e. the centre of the first FE from the surface at $x=0$, and in the centre of the simulation box from top to bottom. (a), (e) and (i) are the stress fields for the infinite-domain solution, $\tns{\hat{\sigma}}$; (b), (f) and (j) are those obtaine from analytic tractions + FEM, $\tns{\hat{\sigma}}^{\textrm{A}}$; (c), (g) and (k) are those obtained using numeric tractions ($Q = 1$) + FEM, $\tns{\hat{\sigma}}^{\textrm{N}}$. (a) to (c) represent the $xx$; (e) to (g) the $yy$; and (i) to (k) the $xy$ components of the stress tensor.}
    \label{f:head_vs_ana_vs_num_epar}
\end{figure}
\Cref{f:head_vs_ana_vs_num_eperp} shows the stress fields corresponding to analytic expressions for image stress components, $\hat{\sigma}_{ij}$ in \cref{eq:imageStressAnalyticEdge1} where no superscript denotes the infinite-domain singular expressions, the $\textrm{A}$ superscript are the stresses calculated from analytic tractions and the $\textrm{N}$ superscript are those coming from numeric tractions where $Q = 1$ (same nomenclature in \cref{f:head_vs_ana_vs_num_epar,f:head_vs_ana_vs_num_screw}). The setup corresponds to the one described in \cref{f:headvstractionfem} where $\vec{b} = \vec{b}_{\textrm{e1}} = [1\, 0\, 0]$ and the dislocation is found at $(26, 1000)$.

@@

    \subfloat[%$\hat{\sigma}_{yz}$.
        \hspace{-0.8cm}\label{sf:szyscrew}]
    {
        {\includegraphics[width=0.25\linewidth]{syzscrew_leg.pdf}}
    }~
    \subfloat[%$\hat{\sigma}^\textrm{A}_{yz}$.
        \hspace{-0.8cm}\label{sf:syzAscrew}]
    {
        {\includegraphics[width=0.25\linewidth]{syzAscrew_leg.pdf}}
    }~
    \subfloat[%$\hat{\sigma}^\textrm{N}_{yz}$.
        \hspace{-0.8cm}\label{sf:syzNscrew}]
    {
        {\includegraphics[width=0.25\linewidth]{syzNscrew_leg}}
    }~
    \subfloat{
        \includegraphics[height = 0.227\linewidth]{syzNAscrew_leg.pdf}
    }
    \caption{Image stresses for a screw dislocation with $\vec{l} = [0\, 0\, 1]$, $\vec{b} = [0\, 0\, 1]$, where $a = 5$, with coordinates $(26,\, 1000)$, i.e. the centre of the first FE from the surface at $x=0$, and in the centre of the simulation box from top to bottom.  (a) and (e) are the stress fields for the infinite-domain solution, $\tns{\hat{\sigma}}$; (b) and (f) are those obtaine from analytic tractions + FEM, $\tns{\hat{\sigma}}^{\textrm{A}}$; (c) and (g) are those obtained using numeric tractions ($Q = 1$) + FEM, $\tns{\hat{\sigma}}^{\textrm{N}}$. (a) to (c) represent the $xz$; (e) to (g) the $yz$.}
    \label{f:head_vs_ana_vs_num_screw}
\end{figure}
Things get markedly more interesting when looking at $\vec{b} = \vec{b}_{\textrm{e2}} = [0\, 1\, 0]$ in \cref{f:head_vs_ana_vs_num_epar} for a dislocation in the same place, $(26, 1000)$. Of particular note is $\hat{\sigma}_{xx}$, where a comparison between \cref{sf:sxxepar,sf:sxxAepar} and \cref{sf:sxxNepar} reveals one of the major issues with numeric tractions. If we look at the neighbourhood of the dislocation (just to the right), we will find a sign inversion i.e. yellow and green contours as opposed to purple and blue, as well as a drastically different isosurface shape. Image stresses like those can lead dislocations to behave quite differently than they should, particularly if the sign inversion also has a significantly different magnitude than the correct solution. Specifically, there is a region in the positive $x$-direction away from the dislocation where $\hat{\sigma}_{xx}$ is tensile rather than compressive. It is as absurd as a ship that floats by lowering the water level.

@@

    \subfloat[$\hat{\sigma}_{xz}$.\label{sf:line_sxzscrew}]
    {
        {\includegraphics[width=0.3\linewidth]{line_sxzscrew_leg.pdf}}
    }~
    \subfloat[$\hat{\sigma}_{yz}$.\label{sf:line_syzscrew}]
    {
        {\includegraphics[width=0.3\linewidth]{line_syzscrew_leg.pdf}}
    }
    \caption{Line plots of the image stresses of a dislocation at $(26, 1000)$ for a line going through $x = 103$ (start of the third element) for the analytic image stresses, as well as those calculated with numeric and analytic tractions. (a) to (c) correspond to $\vec{b} = \vec{b}_{\textrm{e1}}$, (d) to (f) $\vec{b} = \vec{b}_{\textrm{e2}}$ and (g) to (h) to $\vec{b} = \vec{b}_{\textrm{s}}$.}
    \label{f:line_head_vs_ana_vs_num}
\end{figure}
From \cref{f:head_vs_ana_vs_num_eperp,f:head_vs_ana_vs_num_epar,f:head_vs_ana_vs_num_screw} it seems like both analytic and numeric tractions are appropriate in most cases. At least at these scales, there is only one instance where numeric tractions yield very incorrect results. That said, these can have a significant impact on simulations, particularly those where multiple dislocations interact with surfaces in proximity with one another.

@@

From \cref{f:head_vs_ana_vs_num_epar2to11,f:line_head_vs_ana_vs_num_epar2to11} it can be observed that both analytic and numeric tractions converge to similar shapes to each other as expected. However, it also becomes clear that the infinite-domain solution is not totally accurate for image stresses in finite domains due to edge effects.

Lastly, to show how much numeric tractions can affect a result, \cref{f:simulation} shows a few snapshots of the simulation described in \cref{s:method}. The figures shown are not at equivalent times because the simulations using numeric tractions ran indefenitely. The line tension equilibrated with the image forces on the traction-free surface, preventing the dislocation from exiting and keeping it oscillating in a local energy minimum. The simulation using analytic tractions ended when the dislocation completely exited the surface as expected. It is feasable that at some point, the simulation using numeric tractions could jump out of the local minimum due to some spike in the tractions, but the last image was taken for a simulation time approximately 20 times greater than it took the one using analytic tractions to end when the dislocation fully exited the box.
\begin{figure}
    \subfloat[Analytic tractions $t = 0$.]
    {
        {\includegraphics[width=0.3\linewidth]{analytic_0.pdf}}
    }~
    \subfloat[Analytic tractions mid-simulation.]
    {
        {\includegraphics[width=0.3\linewidth]{analytic_50.pdf}}
    }~
    \subfloat[Analytic tractions end.]
    {
        {\includegraphics[width=0.3\linewidth]{analytic_66.pdf}}
    }

    \subfloat[Numeric tractions $t = 0$.]
    {
        {\includegraphics[width=0.3\linewidth]{numeric_0.pdf}}
    }~
    \subfloat[Numeric tractions mid-simulation.]
    {
        {\includegraphics[width=0.3\linewidth]{numeric_78.pdf}}
    }~
    \subfloat[Numeric tractions $t \to \infty$.]
    {
        {\includegraphics[width=0.3\linewidth]{numeric_156.pdf}}
    }
    \caption{Progression of simulations using both traction calculation methods.}
    \label{f:simulation}
\end{figure}

@@

Often the effects of tractions on a simulation can be quite subtle if things do not go catastrophically wrong, but a quick and easy way of seeing how the numeric tractions have a tendency to spike and invert sign (as in \cref{sf:line_sxxepar}) is to pause a simulation while its running and scatter plot the numeric tractions vs analytic ones. The plot will show a positive correlation---as the sign inversions in numeric tractions are relatively rare events---but whichever axis corresponds to the numeric tractions will have the largest range. Using proportional axes makes the differences quite evident even at a quick glance as the aspect ratio will be far from 1:1.

In other parts of our work, we have noted that despite the analytic tractions taking approximately 10 times longer to compute than numerically calculated tractions with $Q = 1$ (which agrees with the findings in \cite{Queyreau}), yield more stable and ultimately faster simulations. Even simple simulations are typically faster when using analytic tractions over numeric ones. This was not the case with the simulation we show here, but the numeric tractions led to a hung simulation, which we have found is quite common. The margin by which using analytic tractions leads to faster overall simulations grows as simulation complexity increases. Only the simplest simulations initially benefit from numeric tractions, but this often decreases and reverses as the simulation advances. Another fortunate side effect of more accurate tractions is the fact that fewer dislocation segments are generated during simulations, which \citet{bromage2018calculating} found when correctly accounting for the displacements generated by dislocations.

Given our findings, we cannot recommend using numeric tractions when analytic ones are available. The losses in computational speed that result from moving from one to the other are more than made up for by the fewer topological operations, fewer generated segments, and more accurate velocities. All of which result in fewer calculations of segment-segment interactions (remote forces), fewer collisions, larger timesteps and ultimately cleaner simulations that run into fewer snags along the way.

There is a case however, for combining both approaches in larger scale simulations. Since both numeric and analytic tractions converge to the same value as the dislocation-surface distance increases, a hybrid approach could possibly be undertaken without many negatives. Furthermore, \cite{Queyreau} also derived a Taylor series expansion of the solutions which can be used in cases where the dislocation-surface distance is large. These may be worth exploring in larger scale simulations. However, in the types of systems we model, other parts of our model tend to be much more rate-limiting than tractions. As such, we have all but ceased to use numeric tractions in our work.
